### 2022-11-28 data/v1b-analysis.rmd
Update images

---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(lubridate)
events <- read_csv("v1b-data.csv") %>%
  mutate(time = as_datetime(time), task = as.integer(task)) %>%
  arrange(time)
```

```{r message=FALSE}
event_names <- events %>% distinct(event_name)

all_participants <- events %>%
  distinct(participant_id) %>%
  .$participant_id

completed <- events %>%
  filter(task == 8, event_name %in% c("complete", "skip")) %>%
  distinct(participant_id) %>%
  filter(! participant_id %in% c()) %>% 
  .$participant_id

starts <- events %>%
  filter(participant_id %in% completed) %>%
  filter(event_name == "start") %>%
  group_by(participant_id, task) %>%
  summarise(start_time = first(time))

ends <- events %>% 
  filter(participant_id %in% completed) %>%
  filter(event_name %in% c("complete", "skip")) %>%
  group_by(participant_id, task) %>%
  summarise(finish_time = last(time))

durations <- events %>%
  filter(participant_id %in% completed) %>%
  left_join(starts, by=c("participant_id", 'task')) %>%
  left_join(ends, by=c("participant_id", 'task')) %>%
  mutate(duration = as.numeric(finish_time - start_time))

durations |> group_by(task) |> summarise(
  avg_time=mean(duration),
  med_time= median(duration),
  min_time = min(duration),
  max_time=max(duration),
  sd_time = sd(duration))

attemps <- events %>%
  filter(participant_id %in% completed) %>% 
  filter(event_name == "attempts")

```

```{r}

err_left <- function (n, m, s) {
  return (m - qt(0.975,df=n-1)*s/sqrt(n))
}

err_right <- function (n, m, s) {
  return (m + qt(0.975,df=n-1)*s/sqrt(n))
}

errors <- durations |>
  group_by(task, treatment) %>%
  summarise(
    l = err_left(n(), mean(duration), sd(duration)),
    m = mean(duration),
    u = err_right(n(), mean(duration), sd(duration))
  )

durations |>
  ggplot() +
  geom_bar(
    durations,
    mapping = aes(x = factor(task),
                  y = duration,
                  fill = treatment),
    position = "dodge",
    stat = "summary",
    fun = "mean"
  ) +
  geom_errorbar(
    data = errors,
    mapping = aes(
      x = factor(task),
      y = m,
      ymin = l,
      ymax = u,
      group = treatment
    ),
    position = position_dodge(0.9),
    alpha=0.5,
    width=0.6
  ) +
  labs(x = "Tasks", y = "Time (Seconds)", fill="Debugging Tool:", title="User study 2") + 
  scale_x_discrete(expand = c(0,0))  +
  scale_y_continuous(expand = c(0,0)) + 
  scale_fill_manual(
    values=c("#8fbfd5", "#b14863"),
    labels=\(inputs) map_chr(inputs, ~ ifelse(. == "Chameleon", "ChameleonIDE", "GHC"))
    ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.border = element_blank(),
    axis.line.y = element_line(colour = "black"),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(size = 15),
    axis.title = element_text(size = 14),
    text = element_text(size = 14)
    )

ggsave("r2-data.pdf", device = "pdf", plot = last_plot(), units ="mm", width = 210)

```

```{r}
d = durations |>
  select(participant_id, treatment, duration) |>
  group_by(participant_id, treatment) |> 
  summarise(duration=mean(duration)) |>
  pivot_wider(names_from = treatment, values_from = duration)

r <- wilcox.test(d$Chameleon, d$Ghc, paired=TRUE)
r
```

```{r}
one.way <- aov(duration ~ treatment, durations)
summary(one.way)
```

```{r}
events |> 
  group_by(participant_id) |>
  summarise(runtime = max(time) - min(time)) |>
  summarise(
    mean_runtime = mean(runtime),
    median_runtime = median(runtime)
    )
```

```{r}
events |>
  ggplot(mapping=aes(x=time)) +
  geom_density()

```