### 2022-12-03 sections/conclusion.tex

\section{Conclusion}

We present \chameleon{}, a type debugging tool for the Haskell programming language. \chameleon{} facilitates a constraint-based type inference engine that provides accurate and comprehensive type error reports. \chameleon{} also features an interactive debugging interface with three different debugging modes, enabling programmers to navigate deep type relations and triangulate conflicting statements in source code. Four experiments with human participants were conducted to evaluate the tool's design with programmers with varied experiences. We found that, particularly for more complex tasks, \chameleon{} helped programmers to fix type errors more quickly than traditional text-based error messages.  Further, programmers actively using \chameleon{} interactive features are shown to fix type errors faster than simply reading the type error output.