### 2022-12-12 sections/conclusion.tex

\section{Conclusion}

We present \chameleon{}, a type debugging tool for the Haskell programming language. Its constraint-based type inference engine provides unbiased and comprehensive error location reporting. \chameleon{} also features an interactive debugging interface with three debugging modes, enabling programmers to interactively explore the stepwise explanations. Our studies evaluated the tool's design with programmers. We found that, particularly for more complex tasks, \chameleon{} helped programmers to fix type errors more quickly than traditional text-based error messages. Further, programmers actively using \chameleon{} interactive features are shown to fix type errors faster than simply reading the type error output.